---
title: "OMI - 2 condition anlaysis kit"
output: html_notebook
---

##This is code to generate:##
##plots with statistics for comparing 2 data sets together##
##dimension reduction plots##
##machine learning plots##

##Data must be located on desktop with a colum titled "Group" that has labels of either "AC" or "QC" and a column called "celltype" that has values equal to "cyto"##

```{r}
##load data and select A and Q sets##
library(gtools)
library(ggplot2)
library(ggsci)
library(tidyverse)
rm(list=ls())
filename <- "allfinaloutputqNSCaNSC masterjust8.csv"
NSCdata_rawtitle <- paste('C:/Users/cmorrow5/Desktop/',filename, sep = "")
NSCdata_raw <- read.csv(NSCdata_rawtitle)
NSCdatacyto <- subset(NSCdata_raw, celltype=="cyto")
mygroups <- c("AC","QC")
NSCdatacontrol <- subset(NSCdatacyto, Group%in%mygroups)
NSCdatacontrol$Group <- factor(NSCdatacontrol$Group, levels=c("AC","QC"))
NSCdata <- NSCdatacontrol

colorA <- "darkblue" 
colorQ <- "red" 
plotcolors <- c(colorA, colorQ)
plotcolors

##normalize data set##
myvars <- c("n.p.mean", "f.p.mean", "n.a1.mean", "n.t1.mean", "n.t2.mean", "f.a1.mean", "f.t1.mean", "f.t2.mean", "Group")
NSCdata_reduc <- NSCdatacontrol[myvars]
last <- ncol(NSCdata_reduc)
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
NSCdata_reduc_norm <- range01(NSCdata_reduc[,-last])
Group <- NSCdatacontrol$Group
NSCdata_reduc_norm_Group <- data.frame(NSCdata_reduc_norm, Group)
set.seed(1)
```

filename <- "042219_flim_outputs_cyto.csv"
filename
NSCdata_raw <- read.csv('C:/Users/cmorrow5/Desktop/',filename,')
NSCdata_raw


##Number of cells/group##
```{r}
ftable(NSCdatacontrol$Group)
```
##NAD(P)H Intensity##
```{r}
setwd("C:/Users/cmorrow5/Desktop/")
pdf("1.pdf")
par(oma=c(3,0.5,0.5,0.5))

fit <- glm(n.p.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=n.p.mean, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("NAD(P)H Intensity"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(1.5), label=Pval), size=10, color="red")

dev.off()

```


##FAD Intensity##
```{r}
setwd("C:/Users/cmorrow5/Desktop/")
pdf("2.pdf")
par(oma=c(3,0.5,0.5,0.5))

fit <- glm(f.p.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=f.p.mean, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("FAD Intensity"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(1.5), label=Pval), size=10, color="red")

dev.off()
```



Redox ratio all cells
```{r}


fit <- glm(rr.mean~ Group, data=NSCdata)
graphLabelsQC <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=rr.mean, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("Redox Ratio"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(0.9), label=Pval), size=10, color="red")


```

NAD Lifetime
```{r}

setwd("C:/Users/cmorrow5/Desktop/")
pdf("3.pdf")
par(oma=c(3,0.5,0.5,0.5))

fit <- glm(n.tm.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=n.tm.mean/1000, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("NAD(P)H  " *tau[m]* " (ns)"))+
  geom_text(data=graphLabelsQC, aes(x=Group, y = c(0.085), label=Pval), size=10, color="red")

dev.off()
```


```{r}

setwd("C:/Users/cmorrow5/Desktop/")
pdf("4.pdf")
par(oma=c(3,0.5,0.5,0.5))

fit <- glm(n.a1.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=n.a1.mean/1000, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("NAD(P)H  " *alpha[1]* " (ns)"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(0.085), label=Pval), size=10, color="red")

dev.off()
```


```{r}
fit <- glm(n.t1.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=n.t1.mean/1000, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("NAD(P)H  " *tau[1]* " (ns)"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(0.75), label=Pval), size=10, color="red")
```
```{r}
fit <- glm(n.t2.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=n.t2.mean/1000, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("NAD(P)H  " *tau[2]* " (ns)"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(3.2), label=Pval), size=10, color="red")
```

FAD Lifetime
```{r}

setwd("C:/Users/cmorrow5/Desktop/")
pdf("5.pdf")
par(oma=c(3,0.5,0.5,0.5))

fit <- glm(f.tm.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=f.tm.mean/1000, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("FAD  " *tau[m]* " (ns)"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(0.1), label=Pval), size=10, color="red")

dev.off()
```


```{r}

setwd("C:/Users/cmorrow5/Desktop/")
pdf("6.pdf")
par(oma=c(3,0.5,0.5,0.5))

fit <- glm(f.a1.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=f.a1.mean/1000, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
   scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("FAD  " *alpha[1]* " (ns)"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(0.1), label=Pval), size=10, color="red")

dev.off()
```


```{r}
fit <- glm(f.t1.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
ggplot(data=NSCdata, aes(x=Group, y=f.t1.mean/1000, color=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
   scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("FAD  " *tau[1]* " (ns)"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(0.5), label=Pval), size=10, color="red")
```


```{r}
fit <- glm(f.t2.mean~ Group, data=NSCdata)
graphLabels <- data.frame(Group = c("QC"),Pval=c(stars.pval(summary(fit)$coefficients[2,4])))
par(pty="s")
ggplot(data=NSCdata, aes(x=Group, y=f.t2.mean/1000, col=as.factor(Group)))+geom_violin(size=1)+
  stat_summary(fun.data="mean_cl_boot", size=0.75)+
  theme_classic(base_size = 28, base_family = "")+
  scale_color_manual(values=plotcolors)+
  theme(legend.position = "none")+
  ylab(bquote("FAD  " *tau[2]* " (ns)"))+
  geom_text(data=graphLabels, aes(x=Group, y = c(2.5), label=Pval), size=10, color="red")
```
pca <- prcomp(NSCdata_reduc_norm_Group[,-last], scale=TRUE)
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
barplot(pca.var.per, main="Scree Plot", xlab="Principal Component", ylab="Percent Variation")
par(mar=c(5,7,1,1), mgp=c(3,1,0))
plot(pca$x[,1], pca$x[,2], pch=16, xlab="PC1", ylab="PC2", "cex.axis"=1, las=1, col=c(colorA, colorQ)[as.numeric(NSCdata_reduc_norm_Group$Group)], cex.lab=2, cex.axis=2)
 legend("bottomleft", legend=c("aNSC", "qNSC"), pch=16, col=plotcolors, lwd=4, seg.len=0, bty="n") 

##PCA##
```{r}

pca <- prcomp(NSCdata_reduc_norm_Group[,-last], scale=TRUE)
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
barplot(pca.var.per, main="Scree Plot", xlab="Principal Component", ylab="Percent Variation")
par(mar=c(5,7,1,1), mgp=c(3,1,0), pty="s")
plot(pca$x[,1], pca$x[,2], pch=16, xlab="Principal Component 1", ylab="Principal Component 2", "cex.axis"=1, las=1, col=c(colorA, colorQ)[as.numeric(NSCdata_reduc_norm_Group$Group)], cex.lab=2, cex.axis=2)
 legend(x=-5.4, y=-3.1, legend=c("aNSC", "qNSC"), pch=16, col=plotcolors, lwd=4, seg.len=0, bty="n", cex=1.5, pt.cex=1) 
```

##tSNE##
```{r}
library(Rtsne)
set.seed(1)
tsne <- Rtsne(NSCdata_reduc_norm_Group[,-last], dims=2, perplexity = 20, verbose=TRUE,max_iter=1000)
colors = rainbow(length(unique(NSCdata_reduc_norm_Group$Group)))
names(colors) = unique(NSCdata_reduc_norm_Group$Group)
par(mar=c(5,7,1,1), mgp=c(4,1,0), pty="s")
plot(tsne$Y[,1], tsne$Y[,2], pch=16, xlab="tSNE dimension 1", ylab="tSNE dimension 2", "cex.lab"=2, "cex.axis"=2, las=1, col=c(colorA, colorQ)[as.numeric(NSCdata_reduc_norm_Group$Group)])
legend(x=7, y=-19, legend=c("aNSC", "qNSC"), pch=16, col=plotcolors, lwd=4, seg.len=0, bty="n", cex=1.75, pt.cex=1) 
```


##UMAP##
```{r}
library(umap)
set.seed(1)
umap <- umap(NSCdata_reduc_norm_Group[,-last])

colors = rainbow(length(unique(NSCdata_reduc_norm_Group$Group)))
names(colors) = unique(NSCdata_reduc_norm_Group$Group)
par(mar=c(5,7,1,1), mgp=c(3,1,0), pty="s")
plot(umap$layout[,1], umap$layout[,2], pch=16, xlab="UMAP dimension 1", ylab="UMAP dimension 2", "cex.lab"=2, "cex.axis"=2, las=1, col=c(colorA, colorQ)[as.numeric(NSCdata_reduc_norm_Group$Group)])
legend(x=1.7, y=-4.2, legend=c("aNSC", "qNSC"), pch=16, col=plotcolors, lwd=4, seg.len=0, bty="n", cex=1.75, pt.cex=1) 

```



